---
title: "STA303 Assignment 2 1001569732"
author:
- "Ted Si Yuan Cheng"
- "Student ID: 1001569732"
date: '2018-02-04'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Question 1
```{r, echo=F}
#Import data
bbw <- read.csv("~/Dropbox/School/3rd Year/Sem 2/STA303/Assignments/2/bbw.csv")
attach(bbw)

#Create variables
smoke <- bbw$smoke
gestation <- bbw$gestation
bwt <- bbw$bwt

#Convert to factors
maturity=array(0,length(gestation))
MatSmoke=array(0,length(smoke))
for (i in 1:length(gestation))
{
  if (gestation[i]<259)
  {maturity[i]=1}
  else if (gestation[i]>293)
  {maturity[i]=3}
  else {maturity[i]=2}
}
for (i in 1:length(smoke))
{
  if (maturity[i]==1 & smoke[i]==1)
  {MatSmoke[i]="PreSmoke"}
  else if (maturity[i]==1 & smoke[i]==0)
  {MatSmoke[i]="PreNoSmoke"}
  else if (maturity[i]==2 & smoke[i]==1)
  {MatSmoke[i]="NorSmoke"}
  else if (maturity[i]==2 & smoke[i]==0)
  {MatSmoke[i]="NorNoSmoke"}
  else if (maturity[i]==3 & smoke[i]==1)
  {MatSmoke[i]="PostSmoke"}
  else {MatSmoke[i]="PostNoSmoke"}
}

maturity <- as.factor(maturity)

```

##Boxplot of birth weight between mothers who smoked and didn't smoke
```{r, echo=F}
#Boxplot for smokers vs non-smokers
groupsmoke <- bwt[smoke==1]
groupnosmoke <- bwt[smoke==0]
boxplot(groupsmoke, groupnosmoke, main = 'Smokers vs Non-smokers 9732')

```

##Boxplot of birth weight among three maturity levels 9732
```{r, echo=F}
#Boxplot for maturity levels
library(ggplot2)
ggplot(bbw, aes(x=maturity, y=bwt, fill=maturity))+geom_boxplot()

```

Body weight seems to increase proportionally to maturity levels

###Boxplot of birth weight among babies grouped by combination of their maturity level and maternal smoking status 9732

```{r, echo=F}
#Boxplot for maturity level and smoking status
ggplot(bbw, aes(x=MatSmoke, y=bwt,fill=MatSmoke))+geom_boxplot()

```

The non-smoker's birth weights appear to higher than their respective group of smokers. Pre-pregnancy smoking seems to cause the lowest birth weight

#Question 2
```{r,echo=F, results='hide'}
#T test for smokers vs nonsmokers
t.test(groupsmoke, groupnosmoke)

```

The true difference in mean between the smokers and non smokers are not equal to 0 (p > 0.05), so we reject $H_0$ = 0 which states there is no difference in the mean birth weight between babies born to smoking versus nonsmoking mothers.

#Question 3

```{r, echo=F, results='hide'}
#ANOVA and Bonferroni test
aov_m <- aov(bwt~maturity)
summary(aov_m)
posthocm <- pairwise.t.test(bwt, maturity, p.adjust.method = 'bonf')
posthocm

```

The one-way ANOVA of body weight versus gestation shows that there is a difference as p < 0.05.

The post-hoc Bonferroni test shows there is a difference among the mean levels of maturity between all level as all values of p < 0.05.

#Question 4

```{r, echo=F, results='hide'}
#ANOVA and Bonferroni test
aov_ms <- aov(bwt~MatSmoke)
summary(aov_ms)
posthocms <- pairwise.t.test(bwt, MatSmoke, p.adjust.method = 'bonf')
posthocms

```

The one-way ANOVA of body weight versus maturity level and smoking status shows that there is a difference as p < 0.05.

The post-hoc Bonferroni test shows that there is a difference between NorSmoke~NorNoSmoke, PreNoSmoke~NorNoSmoke, PreSmoke~NorNoSmoke, PostNoSmoke~NorSmoke, PostSmoke~NorSmoke, PreNoSmoke~NorSmoke, PreNoSmoke~PostNoSmoke, PreSmoke~PostNoSmoke, PreNoSmoke~PostSmoke, PreSmoke~PostSmoke, PreSmoke~PreNoSmoke, as all p<0.05.

#Question 5

```{r, echo=F}
#Check Bonferroni test assumptions
fitms <- lm(bwt~MatSmoke, data=bbw)
par(mfrow=c(1,2))
plot(fitms, which=1:2)

```

```{r, echo=F, results='hide'}
chisq.test(bwt, MatSmoke, simulate.p.value = T)

```

The assumption of normality for the ANOVA test is met from the linear model qq plot showing normality. The assumption of homogeneous variance also seems to hold, as the spreads are quite close together in the residuals vs fitted graph. However, the test for independency is not as the Chi-Squared test showed p = 0.0004998, rejecting the null hypothesis that the baby weight among the six categories of maturity level and smoking status are independent.

The assumptions are met for the Bonferroni method as it can be used more generally than the Tukey. 

#Question 6
##a)
The number of predictor variables would be more than the model in quesiton 4 as two-way anova will require two predictor variables. 

##b)
Yes it would as from question 4 we see that many of the interactions between smokers and non-smokers at various stages of pregnancy show a significant difference.

#Question 7

No, as the Bonferroni method can be used with unequal sample sizes as well as equal sample sizes. Outside of the Bonferroni, this may affect our ability to do a t test for the mean difference of one level of bodyweight for the combination of MatSmoke to the other level, likewise for maturity. Doing a t.test with unequal sample sizes could result in inflated errors if the variances are not equal as well.

#Question 8

Gestation as a quantitative explanatory variable will yield the equation $y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \epsilon_i$.
$y_i$ denotes the birth weight of the $i^{th}$ baby, $x_{i1}$ is gestation age of the $i^{th}$ baby, and $x_{i2}$ is the categorical variabe. The dummy variable will be 1 if the mother smoked, or 0 if she did not. The mean response function is $\mu_{Y0} = \beta_0 + \beta_1 x_{i1}$ for mothers that did not smoke, and $\mu_{Y1} = (\beta_0 + \beta_2) + \beta_1 x_1$ for mothers that did smoke.

Gestation as a factor will yield the equation $y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \epsilon_i$. $y_i$ denotes the birth weight of the $i^{th}$ baby, $x_{i1}$ is the maturity level based on the gestation age of the $i^{th}$ baby. The dummy variable $x_{i1}$ is 1 if the gestation age is less than 259 days , 3 if it is >293, and 2 otherwise. $x_{i2}$ is the same categorical variable it was as the previous equation with the same levels. The mean response functions from this model will be in order of maturity level from 1 to 3 and nonsmokers are  $\mu_{Y1} = \beta_0 + \beta_1$, $\mu_{Y2} = \beta_0 + 2\beta_1$, $\mu_{Y3} = \beta_0 + 3\beta_1$. For mothers who did smoke in order of maturity level 1 to 3 they will be $\mu_{Y1} = \beta_0 + \beta_1 + \beta_2$, $\mu_{Y2} = (\beta_0 + \beta_2) + 2\beta_1$, $\mu_{Y3} = (\beta_0 + \beta_2) + 3\beta_1$.

##Question 9
One possible factor affecting birth weight could be the mother's nutrition intake during pregnancy. The possible levels could be 1 if the mother undereats, 2 if the mother eats a moderate amount, and 3 if the mother overeats. Hypothetically, a mother who undereats will have infants that will be underweight. 

Another factor would be the age of the mother during pregnancy. Studies have shown that mothers who are older than 35 have larger babies, and teens with babies that are underweight. We can therefore have 3 levels as well, 1 being the age of any women/girl under the age of 18, 2 being the range of the age of women from 18 to 35, and 3 being the age of any women 35 and older.

# Appendix
``` {r, ref.label=knitr::all_labels(), tidy=TRUE, echo=TRUE, eval=TRUE, documentation = 1}
```